---
title: "R4BD Assignment"
author: "Nelson López"
output: github_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```
```{r}
library(rsconnect)
rsconnect::deployApp("C:/users/nosle/Desktop/Assigmments/R Assignment/")
```

```{r}
load("C:/users/nosle/Desktop/Assigmments/R Assignment/AirBnB.RData")
```

## Libraries import
For this analysis we will use the following libraries:

```{r}
library("shiny")
library("dplyr")
library("ggplot2")
library("lubridate")
```


## Initial analysis Requirements by the professor 

Relationship between prices and apartment features
Number of apartments per owner
Renting price per city quarter (“arrondissements”)
Visit frequency of the different quarters according to time

## Approach

1- The first look at the data:

- what information do we have?
- what information is missing?
- discover general facts: time period, total number of listings in the dataset

2- Initial data preparation:

- remove irrelevant information;
- reformat the information and imputing missing values;

3- High level analysis:

- Find out answers to the questions above
- Fin out general analysis about the data found

## First Look at the data, What information do we have?

Data provided, consists of 2 tables:

- L Table: Summary information on listing in Paris with general information such as: host information, location, zip code, price, etc.
- R Table: List of listing reservations completed with date 

```{r}
colnames(L)
```
```{r}
colnames(R)
```

## Removing irrelevant information for our analysis

For our analysis we will use few columns and will drop other columns in the cleaning data process:
id, host_name, property_type, bathrooms, bed_type, price, calculated_host_listings_count, listing_url, host_id, host_is_superhost, zip code, country_code, longitude, room_type, bedrooms, weekly_price, neighbourhood, city, beds, monthly_price, latitude, minimum_nights.

## Data Handling and cleaning for answer questions

Importing dplyr library for data manipulation and ggplot to visualizations. Building a new table with our working data

```{r}
listing_df <- select(L, id, host_name, property_type, bathrooms, bed_type, price, calculated_host_listings_count, listing_url, host_id, host_is_superhost, zipcode, country_code, longitude, room_type, bedrooms, weekly_price, neighbourhood, city, beds, monthly_price, latitude, minimum_nights)

```


Checking NA values and if we need to impute them

```{r}
table(is.na(listing_df))
```
Checking NA values per column

```{r}
colSums(is.na(listing_df))
```

We suppress all reviews columns and square feet too because there were 50000 missing values of 52725 obs, this means that we cannot handle this quantity of missing values because they are too many according to the total data. About bedrooms, beds, and bathrooms missing data, we will impute them replacing NA values by the mean because they represents a small percentage of the total data


```{r}
for (j in 1:ncol(listing_df)){
  sel = is.na(listing_df[,j])
  listing_df[sel,j] = mean(listing_df[,j],na.rm=TRUE) 
}
```
Checking data types

```{r}
str(listing_df)
```

Checking zipcode content, after analyse neighbourhood content, i decided to generate location of the listings by zip code. 

```{r}
summary(listing_df$zipcode)
```

## Number of apartments per owner answer

Creating new df to answer "Number of apartments per owner" question, and plotting a sample of 100 observations

```{r}
listing_df$host_id <- factor(listing_df$host_id)
```
```{r}
host_apartment_count <- listing_df %>%  group_by(listing_df$host_id) %>% count(listing_df$host_id)
host_apartment_count$n <- factor(host_apartment_count$n)
host_apartment_count <- host_apartment_count[order(host_apartment_count$n, decreasing = TRUE),]
host_df_toplot <- head(host_apartment_count,100)
host_df_toplot$n <- as.numeric(levels(host_df_toplot$n))[host_df_toplot$n]
host_df_toplot
```
```{r}
df_host_to_plot <- host_apartment_count %>%  group_by(host_apartment_count$n) %>% count(host_apartment_count$n )
df_host_to_plot$n <- as.factor(df_host_to_plot$n)
df_host_to_plot$n <- as.numeric(levels(df_host_to_plot$n))[df_host_to_plot$n]
df_host_to_plot$`host_apartment_count$n` <- as.numeric(levels(df_host_to_plot$`host_apartment_count$n`))[df_host_to_plot$`host_apartment_count$n`]
df_host_to_plot
```
Plotting all observations, we can see that around 41000 host has only 1 accomodation listed, after that we can see some outliers like the host that has more accomodations (155)

```{r}
ggplot(df_host_to_plot, aes(x=`host_apartment_count$n`, y=n) ) +
geom_point(alpha = 2/5) + labs(title= "Number of aparments per host", y="Quantity of Host (Log10)", x = "Quantity of apartment owned", fill = "Legend - Host") + theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), legend.position = "none") + scale_y_continuous(trans = "log10") 
```


### Conclusion 1

After doing some research we see on the plots that, the majority of host has only 1 acommodation published (around 41k hosts), we have some max "top 10" that have more than 60 acommodations. on the Internet we found some people or agencies that work with different properties and manage only the Airbnb acommodation. 
 
## Renting Price per Arrondissemment  

Creating a new df to answer "Renting price per city quarter (“arrondissements”)" question, what are we doing is that we are selecting id, zip code and price and filter by zip code to define arrondissements, In this process I am dropping around 1229 observations that represent around 0.02% of the data and don't add value to our data because those values are not well structured by zip code and to use we have to fix one by one, this decision is taken because we don't want to manipulate few registers in a big data analysis, I decided to drop these observations

```{r}
arrondissement_price_table <- listing_df %>%
  select(c('id', 'host_id','zipcode','price'))%>%
  filter(zipcode %in% c(seq(75001,75020,1),75116))%>%
  mutate(arrondissement=substring(zipcode,5-1,5),price_formatted=(as.numeric(gsub("\\$", "",price))))
```

Imputing NA values in our "price_formatted" column for the mean price

```{r}
colSums(is.na(arrondissement_price_table))
```

```{r}
for (j in 1:ncol(arrondissement_price_table)){
  sel = is.na(arrondissement_price_table[,j])
  arrondissement_price_table[sel,j] = mean(arrondissement_price_table[,j],na.rm=TRUE) 
}
```

Checking data types of arrondissement_price table and computing mean price by arrondissement

```{r}
str(arrondissement_price_table)
```


```{r}
mean_renting_price <- group_by(arrondissement_price_table, arrondissement) %>% summarize(Renting_Price = mean(price_formatted))
```

Plotting Renting price per arrondissement

```{r}
mp_arr_plot <- ggplot(mean_renting_price, aes(x=arrondissement, y=Renting_Price, fill=arrondissement) ) +
geom_bar (stat="identity")+theme_minimal() + labs(title= "Renting Price mean per Arrondissement", y="Renting Price", x = "Arrondissement #", fill = "Arrondissement #") + theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), legend.key.size = unit(0.4, 'cm'), legend.position = "none")
mp_arr_plot
```

### Conclusion 2

As a foreigner, I am surprised that the mean renting price of some arrondissements are more than 100$, the neighbourhood might have a great influence on the listing price as well as its distance to historical monuments.


## Visit frequency of the different quarters according to time answer


"Lubridate" package is to handle date format and subtract year and month from the data, inner join of two tables by listing ID, this way we are getting all listing visits with all info of the L df in this case i utilized arrondissement_price_table that is a cleaned and filtered data.


```{r}
visit_feq_df <- inner_join(R,arrondissement_price_table,by=c('listing_id'='id'))
visit_feq_df$date <- ymd (visit_feq_df$date)
visit_feq_df$Year <- format(as.Date(visit_feq_df$date), "%Y")
visit_feq_df$month <- format(as.Date(visit_feq_df$date), "%m")

```

Group by month year and arrondissement with dplyr and summarize them with respect to number of times by month appears each listing id. 

```{r}
Visit_feq_df_grouped <- visit_feq_df %>% group_by(Year,month,arrondissement) %>% summarize(nummonth=n())
visit_df_to_plot <- Visit_feq_df_grouped %>% filter(arrondissement %in% "12")

```

Plotting visits of Arrondissement 12 vs all years

```{r}
ggplot(visit_df_to_plot, aes(x = month, y = nummonth, group = Year, color = Year, linetype = Year)) +
geom_line()+ xlab("Month of the year") + ylab("Total visits")
```

Plotting visits of all Arrondissements vs all Years

```{r}
ggplot(Visit_feq_df_grouped, aes(x = month, y = nummonth, group = Year, color = Year, linetype = Year)) +
geom_line()+ xlab("Month of the year") + ylab("Total visits") + theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), axis.text.x=element_blank()) +
facet_wrap(~ arrondissement)
```

### Conclusion 3

We can see this in two parts, the first part is about years, we can see easily that there is an increment on complete reservations starting from 2014, it may be because AirBNB started to consolidate its APP as the best "Rent App" and people started to use it as a primary rent app, and the second part is about months, we can see that in all years and all arrondissements there is an increment on complete reservations in June - July and Augusts, correspond to summer in Paris. 

About the arrondissement, we can see that one of the cheapest arrondissements (arrondissement 18) (insight found in the previous analysis) is the most visited one so far, with a considerable number of visits each year.


## Relationship between prices and apartment features

I selected "Property type", "room Type" and "bedrooms" to do the feature vs price analysis, selecting columns to work, computing NA values.

```{r}
pricefeature_df <- select(L, id, property_type, bathrooms, bed_type, price, host_is_superhost, room_type, bedrooms, weekly_price, neighbourhood, city, beds, monthly_price, latitude, minimum_nights)
pricefeature_df$priceformatted <- as.numeric(gsub("\\$", "",pricefeature_df$price))
```

Imputing NA values

```{r}
for (j in 1:ncol(pricefeature_df)){
  sel = is.na(pricefeature_df[,j])
  pricefeature_df[sel,j] = mean(pricefeature_df[,j],na.rm=TRUE) 
}
```

Converting float values of bedroooms to int

```{r}
pricefeature_df$bedrooms <- as.integer(pricefeature_df$bedrooms)
table(pricefeature_df$bedrooms)
```


Relation between Property type vs Price


```{r}
pricefeature_df <- pricefeature_df %>% mutate(property_type = replace(property_type, property_type == "", "Other"))
ggplot(pricefeature_df, aes(x=reorder(property_type, priceformatted), y=priceformatted, fill=property_type) ) +
geom_boxplot(outlier.shape = NA) + coord_flip() + theme(legend.key.size = unit(0.4, 'cm')) +labs(title= "Property Type vs Price analysis", y="Price in $", x = "Features", fill = "Features") + scale_y_continuous(limits = c(0,750))
```


Relation between Room type vs Price

```{r}
ggplot(pricefeature_df, aes(x=reorder(room_type, -priceformatted), y=priceformatted, fill=room_type) ) +
geom_boxplot(outlier.shape = NA) + theme(legend.key.size = unit(0.4, 'cm')) + labs(title= "Room type vs Price analysis", y="Price in $", x = "Features", fill = "Features") + scale_y_continuous(limits = c(0,250))

```

Convertir bedrooms column to factor to plot.

```{r}
pricefeature_df$bedrooms <- as.factor(pricefeature_df$bedrooms)
```


Relation between Bedrooms quantity vs Price


```{r}
ggplot(pricefeature_df, aes(x=bedrooms, y=priceformatted, fill=bedrooms) ) +
geom_boxplot(outlier.shape = NA) + coord_flip() + theme(legend.key.size = unit(0.4, 'cm')) + labs(title= "Bedrooms Quantity vs Price analysis", y="Price in $", x = "Features", fill = "Features") + scale_y_continuous(limits = c(0,1000))

```



### Conclusion 4

I selected "Property type", "room Type" and "bedrooms" to do a cross analysis between prices.
We can see that a shared room is always cheaper than an entire house or apartment. About property type, there is a considerable difference between few types like "Chalet, Villa" that are the most expensive types and the cheapest ones are "Tent", "Cave" and "Dorm", speaking about bedrooms quantity we see a clearly relation between while the accommodation have more bedrooms they are more expensive excepting two situations (7,9).


